---
title: "QTM 150"
subtitle: "ggplot2 <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "Summer"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

You now know:

* The main objects in R.
* How to do basic operations with datasets.
* Numeric variable plots with ggplot2.

**Great job!!**

Do you have any questions?

Today we are going to learn more **ggplot**!
]

---

# This week

.font130[

We will have a **quiz** posted today. Due by **Monday**.

Two chapters of the DataCamp course related to this content has been posted. They are due next week.

Our GitHub page is: https://github.com/umbertomig/qtm150
]

---

# Today's Agenda

.font130[

`ggplot2` graphs: 
* Graphs for numeric variables (recap)
* Graphs for discrete variables
* Graphs for discrete x discrete variables
* Graphs for discrete x numeric variables
* Graphs for numeric x numeric variables
]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/tips.csv') # or your locale folder
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/PErisk.csv') # or your locale folder
head(PErisk, 2)

data(USArrests)
```
]

---
class: inverse, center, middle

# ggplot underlying logic

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# ggplot underlying logic

.font130[

`ggplot` is based on the *grammar of graphs* idea. This idea emphasizes that all graphs are composed of three elements:

* A dataset

* A coordinated system (**mapping** and **aes**)

* Geometric figures (**geom**s)

From these elementary building blocks, we can build any graph we want.

]

---
class: inverse, center, middle

# Graphs for numeric variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# geom_histogram

.font130[

A histogram is a great graph for a numeric variable.

In the dataset USArrests, we can have the histogram for the murder variable:

```
ggplot(data = USArrests) +
  geom_histogram(mapping = aes(x = Murder), bins=15)
```

**Your turn:** Make a histogram of the variable *tip* in the dataset **tips**.

]

---

class: inverse, center, middle

# Graph for discrete variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# geom_bar

.font130[

A bar plot for the *courts* variable in the dataset **PErisk** would tell us how many countries had independent courts in 1992:

```
ggplot(data = PErisk) +
  geom_bar(mapping = aes(x = courts))

PErisk$courts = as.factor(PErisk$courts)
levels(PErisk$courts) <- c('Non-independent', 'Independent')

ggplot(data = PErisk) +
  geom_bar(mapping = aes(x = courts))

ggplot(data = PErisk) +
  geom_bar(aes(factor(prsexp2)), 
           fill = rainbow(6), color = 'black')
```

**Your turn:** Make a barplot of the variable *day* in the dataset **tips**.

]

---

class: inverse, center, middle

# Graphs discrete x discrete variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Mosaic-plots

.font130[

A bar-plot with two variables, one against the other, gives us a good idea whether they are related with each other.

This code plots the variable *corruption* against the variable *courts*:

```
ggplot(data = PErisk) +
  geom_bar(mapping = aes(x = factor(prscorr2), 
                         fill = as.factor(courts)),
           position = 'fill')

PErisk$corruption = ifelse(PErisk$prscorr2>2, 'Low', 'High')
ggplot(PErisk) +
  geom_bar(mapping = aes(x = corruption, fill = as.factor(courts)),
           position = 'fill')
```

**Your turn:** Make a mosaic-plot of the variables *sex* and *smoke*, in the dataset **tips**.

]

---


class: inverse, center, middle

# Graphs discrete x numeric variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Box-plots for multiple categories

.font130[

To visualize the variation in a numeric variable, conditioning by a discrete variable, we can use the `box-plot` (or violin-plots) in a clever way:

```
ggplot(PErisk) +
  geom_boxplot(mapping = aes(x = as.factor(corruption), y = barb2))

ggplot(PErisk) +
  geom_violin(mapping = aes(x = as.factor(corruption), y = barb2))
```

**Your turn:** Plot the variable *tip* against the *time* of the day, in the dataset **tips**.

]

---



class: inverse, center, middle

# Graphs numeric x numeric variables

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Scatter-plots

.font130[

Two numeric variables, one against the other, are better visualized using a scatter-plot:

```
ggplot(data = PErisk) +
  geom_point(mapping = aes(x = barb2, y = gdpw2))
  
ggplot(data = PErisk) +
  geom_point(mapping = aes(x = barb2, y = gdpw2)) +
  geom_smooth(mapping = aes(x = barb2, y = gdpw2), method = 'lm')
```

]

---

# Scatter-plots

.font130[

Two numeric variables, one against the other, are better visualized using a scatter-plot:

```
ggplot(data = PErisk) +
  geom_point(mapping = aes(x = barb2, y = gdpw2, color = courts)) +
  geom_smooth(mapping = aes(x = barb2, y = gdpw2), method = 'lm')
              
ggplot(data = PErisk) +
  geom_point(mapping = aes(x = barb2, y = gdpw2, color = courts)) +
  geom_smooth(mapping = aes(x = barb2, y = gdpw2, color = courts), 
              method = 'lm')
```

**Your turn:** In the dataset **tips**, plot the variable *tip* against the variable *totbill*, with and without differentiating by *smoker*.

]

---


class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>