---
title: "Homework 6 -- QTM 150"
author: "Your name here"
output: html_document
---

## Homework 6

**1** (2 pt). Load the packages `tidyverse`. Load the dataset `voteincome.csv` (check the class GitHub / Canvas). To load the dataset, use the function `read_csv`, from the `tidyverse` library.

```{r}
# Your answers here
```

The dataset has the following informations:

* **VoteIncome**:

>Sample Turnout and Demographic Data from the 2000 Current Population Survey
>
>This data set contains turnout and demographic data from a sample of respondents to the 2000 Current Population Survey (CPS). The states represented are South Carolina and Arkansas. The data represent only a sample and results from this example should not be used in publication.
>
>
>Format
A data frame containing 7 variables ("state", "year", "vote", "income", "education", "age", "female") and 1500 observations.
>
>vote
>an integer vector taking on values "1" (Voted) and "0" (Did Not Vote)
>
>income
>an integer vector ranging from "4" (Less than \$5000) to "17" (Greater than \$75000) denoting family income. See the CPS codebook for more information on variable coding
>
>education
>an integer vector ranging from "1" (Less than High School Education) to "4" (More than a College Education). See the CPS codebook for more information on variable coding
>
>References
>https://www.census.gov/programs-surveys/cps.html
>

**2** (2 pt). Make a boxplot (or violin-plot) of income, segmented by education. What can you infer from this plot?

```{r}
# Your answers here
```

_Answers here_

**3** (2 pt). Are more educated people more likely to vote? Check that by making a mosaic-plot of `vote` against `education`.

```{r}
# Your answers here
```

_Answer here_

**4** (2 pt). Make a barplot of the variable `education`. Plot each bar using a different color.

```{r}
# Your answers here
```

**5** (2 pt). Make a plot to answer the following question: is there a relation between age and income? (Hint: the plot is going to look ugly. Add a trend line to see the relationship.)

```{r}
# Your answers here
```

_Answer here_