---
title: "QTM 150"
subtitle: "Week 13 -- Functions, Loops, and Other Programming <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "Apr 23"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "https://cloud.r-project.org/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>

# Recap

.font130[

You now know:

* The main objects in R.
* How to do basic operations with datasets.
* How to create graphs and plots.
* Data manipulation with `dplyr`.
* Graphs with `ggplot`.

**Great job!!**

Do you have any questions?

Today we are going to learn functions, loops, and elements of statistical programming!
]

---

# This week

.font130[

We will have a **quiz** posted today after 4:00 PM. Due by **Monday**.

We will **not** have a **problem set** this week. Focus on your final projects.

Our GitHub page is: https://github.com/umbertomig/qtm150
]

---

# Final Project

.font130[

The final projects are due by Apr 28 (Wednesday).

Next class we will have our final project presentations. Make sure to come to class. 

Next class is the only class that I require attendance synchronously. Sorry for the folks in different time zones, but this one is important.

Make sure to upload your slides before Friday. Slides are due by Apr 29. I'll this deadline to Canvas later today.

]

---

# Today's Agenda

.font130[

`if` statements

`function` creation

`for` loops

]

---

class: inverse, center, middle

# Getting Started

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Getting Started: loading packages

.font130[
```{r}
# Loading tidyverse
library(tidyverse)
```
]

---

# Loading datasets

.font130[
```{r, tidy=FALSE}
# Loading tips dataset
tips <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/tips.csv') # or your locale folder
head(tips, 2)

# Loading PErisk dataset
PErisk <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/PErisk.csv') # or your locale folder
head(PErisk, 2)

data(USArrests)
```
]

---
class: inverse, center, middle

# if statements

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# if statements

.font130[

In statements control the flow of a code.

They create a condition based on some variable, and execute a piece of code when the condition is `TRUE`.

Example in pseudo-code:

```
if (number > 10) {
  ...execute code for the bigger than 10 case...
} else {
  ...execute code for the smaller than or equals 10 case...
}
```

]

---

# ifelse

.font130[

In R we have a function to conveniently build if-else statements.

If-else consist in binary conditions: do something if true, or something else if false.

Syntax: `ifelse(condition, code-if-true, code-if-false)`

For example:

```
n = 4
ifelse(n>2, n^2, n/5)

n = 1
ifelse(n>2, n^2, n/5)

n = 2
ifelse(n>2, n^2, n/5)
```

]

---

# ifelse

.font130[

More examples:

```
## no need to interact with the condition:
n = 2
ifelse(n>2, 'bigger than 2', 'smaller than or equal 2')

## gender: vector operation
gen <- c(0,0,0,1,1,0,1,1,1,0)
ifelse(gen == 1, 'Fem', 'Masc')

## gender: vector operation
gen <- c(0,0,2,1,1,2,1,1,1,0)
ifelse(gen == 1, 'F', ifelse(gen == 0, 'M', 'Non-binary'))
```

**Your turn**: in the tips dataset, take the variable `day`, and create a condition that shows whether the day is weekday or weekend.

]

---

# "if - else if - else" statements

.font130[

There are statements that are bigger in nature than simple ifelse ones.

Example in pseudo-code:

```
if (number > 10) {
  ...execute code for the bigger than 10 case...
} else if (number > 2 & number <= 10) {
  ...execute code for this condition...
} else {
  ...execute code for otherwise...
}
```

]

---

# "if - else if - else" statements

.font130[

Example:

```{r}
n = 20
if(n==1) {
  print('Number equals 1')
} else if (n==2) {
  print('Number equals 2')
} else if (n==3) {
  print('Number equals 3')
} else if (n==4) {
  print('Number equals 4')
} else if (n<0) {
  print('Number smaller than zero')
} else {
  print('Invalid number!')
}
```

]

---

class: inverse, center, middle

# functions

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# functions

.font130[

Sometimes we need to create our own functions.

Create functions is simple. The syntax is straightforward:

```
name_function <- function(par1, par2, par3, ...) {
  ...code execute...
  return(...return smt...)
}
```

Example in pseudo-code:

```
testn <- function(n) {
  if (number > 10) {
    ...execute code for the bigger than 10 case...
  } else {
    ...execute code for the smaller than or equals 10 case...
  }
  return(...smt...)
}
```

]

---


# functions

.font130[

Example: BMI index function

```{r}
bmi <- function(w, h) {
  bmi = w / (h^2)
  return(bmi)
}
# BMI person 70 kg and 1.75 m
bmi(70, 1.75)
```

**Your turn** create a function that receives weight in pounds and heights in inches and return the BMI.

---

class: inverse, center, middle

# loops

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# loops

.font130[

There are two ways of creating loops: `while` and `for`.

While loops are useful to run a piece of code until a condition is satisfied. For example:

```
i = 1
while(i < 50) {
  print(i)
  print('i is smaller than 50!')
  print(i^10)
  i = i+1
}
```

]


---

# loops

.font130[

Example: infinite loop:

```
while(TRUE) {
  x = readline("Type something or break: ")
  if(x == 'break'){
    break
  }
  cat("You typed: ", x)
}
```

]

---

# loops

.font130[

The `for` loop is a bit different: it receives a counter, and run until the counter is not exhausted.

Example:

```
for (i in 1:10) {
  print(i)
  print(i^5)
}
```

]

---

# loops

.font130[

Example: let's run over the columns of a dataset, and display the class of each variable:

```
for (i in names(tips)) {
  print(class(tips[,i]))
}
```

**Your turn**: Create a function that displays the summary of the numeric or integer variables, and table of the other variables.

]

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>