---
title: "QTM 150"
subtitle: "Week 6 -- Working with Data <html><div style='float:left'></div><hr color='#f2a900' size=1px width=800px></html>"
author: "Umberto Mignozzetti"
date: "Mar 05"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    nature: 
      beforeInit: "https://raw.githubusercontent.com/umbertomig/qtm151/main/slides/wmisc/macros.js"
      highlightStyle: github
      ratio: 4:3
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)
```

<style>

.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 6px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #f2a900;
}

.orange {
  color: #f2a900;
}
</style>
# Recap

.font150[

You now know:

* The main objects in R.
* How to do basic operations with datasets.
* How to create graphs and plots.

**Great job!!**

Do you have any questions?

**Reminder:** The quiz for this class will be posted at 4:00PM. The quiz is due Monday, 11:59 PM.

]

---

# Today's Agenda

.font150[

Today we will learn how to work with data.

* We will repeat some of the things that we did on DataCamp.

* This lab will reinforce the DataCamp learning, while helping you improve some gaps in your knowledge.

* Please ask questions!

]

---

class: inverse, center, middle

# Importing and Saving Data

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

# Importing and Saving Data

.font130[

Importing data from GitHub is easy. Let's do another time in here?

```{r}
gss <- read.csv('https://raw.githubusercontent.com/umbertomig/qtm150/master/datasets/gss.csv')

# Checking dimension
dim(gss) # This dataset contains 53474 rows and 14 variables
```

If GitHub does not work for you, use it in locale: download it to your computer.
]

---

# Importing and Saving Data

.font130[

If the dataset is big, you can always work with a chunk of it to get your coding up and running.

* This has the huge advantage of making things faster.

```{r}
# Random number seed
set.seed(123) # This is not strictly required

# Extracting a chunk
gss100<-gss[sample(nrow(gss),100),] 
dim(gss100) 
```

And we can now save our dataset in our computer.

]

---


# Importing and Saving Data

.font130[

To save the dataset, we need the function `write.csv`. Let's check this function, and save our subsample?!

```{r}
# Check your working directory
getwd() #be sure to check your working directory!

# Save a new dataset in your working directory
write.csv(gss100, "smallgss.csv")
```

We can go ahead and check the folder to see if the dataset is there!

]

---

# Observing our data

.font130[

```{r}
# Print variable names
names(gss100)

# Print dataset dimension
dim(gss100)

# Print summary
summary(gss100)
```

]

---

# Observing our data

.font130[

```{r}
# First observations
head(gss100)

# Last observations
tail(gss100)

# Missing data (very important!)
sum(is.na(gss$age))

# Learning about a function that we have not used before...
?is.na
```

]

---

# Extracting parts

.font130[

Remember that a data.frame has the similar structure as a matrix [rows, columns], and each variable is a vector.

```{r}
#dataset[row, column]
gss100[1,2]

# dataset[rows, columns]
gss100[1:5, c(2,5)] 
```

]

---

# Extracting parts

.font130[

Remember that a data.frame has the similar structure as a matrix [rows, columns], and each variable is a vector.

```{r}
#dataset$variable, the whole column
gss100[,7]
gss100$marital

# dataset$variable[rows]
gss100$marital[1:10]
```

]

---

# R-Base plots
.font130[

We learned `qplot` in the previous class. However, if you need quick (but sadly ugly) plots, you use the R-Base plots.

Here we are going to do a few of these plots.

Histograms:

```{r}
hist(gss100$age)
```

---

# R-Base plots

.font130[

Barplots:

```{r}
barplot(table(gss100$marital))
```

---

# R-Base plots

.font130[

Boxplots (comparing numerical and categorical):

```{r}
boxplot(gss100$age~gss100$sex)
```

---

class: inverse, center, middle

# Questions?

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html> 

---

class: inverse, center, middle

# Have a great weekend!

<html><div style='float:left'></div><hr color='#f2a900' size=1px width=720px></html>